### 2022-03-26 chapters/sctld/article.tex
modif post-defense - part I

Using these three velocity fields, virtual particles were then released on all the reefs composing the FRT to model the dispersal of material carrying the disease causative agent. The locations of the reefs of Florida were extracted from the "coral reefs and hardbottom" layer of the Unified Florida Reef Tract Map \citep{fwc2017unified}. The polygons of this reef map were then further divided into 500 m $\times$ 500 m squares in order to track the propagation of the disease with a finer geographical resolution, generating a total of 16,823 polygons. At the beginning of each simulated month and for each type of current, a total of about $1.5 \times 10^6$ particles were released over all the reef polygons. These particles had a state composed of their polygon of origin as well as their mass, that they lose at a constant rate $\gamma$ as they were moved by surface, mean or bottom currents. The value of $\gamma$ was chosen so that particles had a half-life of 30 days. When the particles were brought over a reef polygon by currents, the amount of disease mass that landed on the polygon was recorded in monthly potential connectivity matrices, whose entries are denoted $C_{ij}$. The matrix rows correspond to the source reefs and the columns correspond to the destination reefs. Hence $C_{ij}$ represents the mass of diseased material originating from sub-reef $i$ that had settled on sub-reef $j$. This matrix was then normalized by dividing each of its rows $i$ by the total mass of particles released on polygon $i$ in order to obtain the normalized potential connectivity matrix $\tilde{C}$, whose entry $\tilde{C}_{ij}$ gives the probability that disease agents produced on sub-reef $i$ settle on sub-reef $j$. Connectivity matrices were computed for each type of current and for each month of the simulated period.

@@

The spread of the SCTLD throughout the FRT was simulated using a connectivity-based \cite{kermack1927contribution} SIR model. SIR models are among the most standard epidemiological models. They divide individuals into three compartments: susceptible ($S$), infectious ($I$) and removed ($R$). When affected by the disease, susceptible individuals become infectious and infect other susceptible individuals until they are removed from the system, either through recovery or death. Such models usually rely on the hypothesis of an homogeneous, well-mixed population. To account for the spatial heterogeneity of the FRT, the basic SIR formulation was modified by considering the fractions of susceptible ($S_j$), infectious ($I_j$) and removed ($R_j$) corals of each sub-reef $j$. Although the pathogen of SCTLD has not been identified, studies suggest a contagious mode of transmission \citep{aeby2019pathogenesis,muller2020spatial}. Here, we use 'infectious' to denote disease agents that could be passed from individual to individual, which are responsible for causing disease signs. We note that the disease agents could be biological or non-biological in nature. In the present study's epidemiological model, individual reefs interact through the exchange of disease agents as represented by the different connectivity matrices. For each sub-reef $j$ and at any time, the following relations hold: $0\leq S_j,I_j,R_j\leq 1$ and $S_j+I_j+R_j=1$. The evolution of these fractions through time is governed by the following equations:
\begin{equation}
    \begin{aligned}
        \dfrac{dS_j}{dt} &= -\beta\sum_i\dfrac{A_i}{A_j}I_i\tilde{C}_{ij}S_j - \beta'(I_j)S_jI_j \\
        \dfrac{dI_j}{dt} &= \beta\sum_i\dfrac{A_i}{A_j}I_i\tilde{C}_{ij}S_j + \beta'(I_j)S_jI_j - \sigma I_j \\
        \dfrac{dR_j}{dt} &= \sigma I_j
    \end{aligned}\label{eq:epidemio}
\end{equation}
where $\tilde{C}_{ij}$ is the entry corresponding to reef pair $(i,j)$ of the normalized potential connectivity matrix [-], $A_i$ is the area of reef polygon $i$ [km$^2$], $\sigma$ is the mortality rate [day$^{-1}$], and $\beta$ and $\beta'(I_j)$ are the inter- and intra-reef disease transmission rates [day$^{-1}$], respectively. In this model, diseased corals of sub-reef $i$ can 'infect' corals of sub-reef $j$ if there is non-zero probability of disease agents exchange from sub-reef $i$ to sub-reef $j$, given by $\tilde{C}_{ij}$. Moreover, to account for coral resistance to the disease, the intra-reef transmission function $\beta'(I_j)$ has the shape of a smooth step function of the fraction of infectious corals $I_j$ and is expressed as follows:
\begin{equation}
    \beta'(I_j) = \dfrac{\beta_0'}{2}(1+\tanh[(I_j-I_0)/\tau]),\label{eq:beta}
\end{equation}
where $I_0$ is a threshold on the infection population above which intra-reef transmission becomes significant, and $\tau$ is a measure of the interval over which the transition from low to high transmission occurs. As long as the fraction of infectious corals on sub-reef $j$ is below $I_0$, the only infection mechanism taking place is connectivity-driven transmission at rate $\beta$. Once the threshold is approached, intra-reef transmission with rate $\beta'_0$ is activated. A larger value of threshold $I_0$ corresponds to a greater resistance of corals to the disease, and therefore a slower spread of the disease within reef $j$. Coral reproduction and natural (\ie~non SCTLD-related) death rates are not taken into account in this model, which amounts to assume that they balance each other. For this study the same values were used for $\beta$ and $\beta'_0$.
%This assumption seems reasonable since coral populations of the FRT were declining prior to the emergence of the SCTLD outbreak.

@@

\begin{figure}
    \centering
    \includegraphics[width=.99\linewidth]{chapters/sctld/figures/hull_example.png}
    \caption{Method used to compute the disease front displacement during a simulated time interval.\textbf{a.} Concave hull $H_0$ of the diseased sub-reefs at the beginning of the simulated period. \textbf{b.} Concave hull $H_1$ of the sub-reefs diseased during the simulated time interval. \textbf{c.} Arrow showing the computed front displacement during simulated time interval between $H_0$ and $H_2$, the union of $H_0$ and $H_1$.}
    \label{fig:hull}
\end{figure}

>>>>

Using these three velocity fields, virtual particles were then released on all the reefs composing the FRT to model the dispersal of material carrying the disease causative agent. The locations of the reefs of Florida were extracted from the "coral reefs and hardbottom" layer of the Unified Florida Reef Tract Map \citep{fwc2017unified}. The polygons of this reef map were then further divided into 500 m $\times$ 500 m squares in order to track the propagation of the disease with a finer geographical resolution, generating a total of 16,823 polygons. At the beginning of each simulated month and for each type of current, a total of about $1.5 \times 10^6$ particles were released over all the \modif{500$\times$500 sub-reef polygons assuming an uniform coral cover}. These particles had a state composed of their polygon of origin as well as their mass, that they lose at a constant rate $\gamma$ as they were moved by surface, mean or bottom currents. The value of $\gamma$ was chosen so that particles had a half-life of 30 days. When the particles were brought over a reef polygon by currents, the amount of disease mass that landed on the polygon was recorded in monthly potential connectivity matrices, whose entries are denoted $C_{ij}$. The matrix rows correspond to the source reefs and the columns correspond to the destination reefs. Hence $C_{ij}$ represents the mass of diseased material originating from sub-reef $i$ that had settled on sub-reef $j$. This matrix was then normalized by dividing each of its rows $i$ by the total mass of particles released on polygon $i$ in order to obtain the normalized potential connectivity matrix $\tilde{C}$, whose entry $\tilde{C}_{ij}$ gives the probability that disease agents produced on sub-reef $i$ settle on sub-reef $j$. Connectivity matrices were computed for each type of current and for each month of the simulated period.

@@

The spread of the SCTLD throughout the FRT was simulated using a connectivity-based \cite{kermack1927contribution} SIR model. SIR models are among the most standard epidemiological models. They divide individuals into three compartments: susceptible ($S$), infectious ($I$) and removed ($R$). When affected by the disease, susceptible individuals become infectious and infect other susceptible individuals until they are removed from the system, either through recovery or death. Such models usually rely on the hypothesis of an homogeneous, well-mixed population. To account for the spatial heterogeneity of the FRT, the basic SIR formulation was modified by considering the fractions of susceptible ($S_j$), infectious ($I_j$) and removed ($R_j$) corals of each sub-reef $j$. Although the pathogen of SCTLD has not been identified, studies suggest a contagious mode of transmission \citep{aeby2019pathogenesis,muller2020spatial}. Here, we use 'infectious' to denote disease agents that could be passed from individual to individual, which are responsible for causing disease signs. We note that the disease agents could be biological or non-biological in nature. In the present study's epidemiological model, individual reefs interact through the exchange of disease agents as represented by the different connectivity matrices. For each sub-reef $j$ and at any time, the following relations hold: $0\leq S_j,I_j,R_j\leq 1$ and $S_j+I_j+R_j=1$. The evolution of these fractions through time is governed by the following equations:
\begin{equation}
    \begin{aligned}
        \dfrac{dS_j}{dt} &= -\beta\sum_i\dfrac{A_i}{A_j}I_i\tilde{C}_{ij}S_j - \beta'(I_j)S_jI_j \\
        \dfrac{dI_j}{dt} &= \beta\sum_i\dfrac{A_i}{A_j}I_i\tilde{C}_{ij}S_j + \beta'(I_j)S_jI_j - \sigma I_j \\
        \dfrac{dR_j}{dt} &= \sigma I_j
    \end{aligned}\label{eq:epidemio}
\end{equation}
where $\tilde{C}_{ij}$ is the entry corresponding to reef pair $(i,j)$ of the normalized potential connectivity matrix [-], $A_i$ is the area of reef polygon $i$ [km$^2$], $\sigma$ is the mortality rate [day$^{-1}$], and $\beta$ and $\beta'(I_j)$ are the inter- and intra-reef disease transmission rates [day$^{-1}$], respectively. In this model, diseased corals of sub-reef $i$ can 'infect' corals of sub-reef $j$ if there is non-zero probability of disease agents exchange from sub-reef $i$ to sub-reef $j$, given by $\tilde{C}_{ij}$ \modif{, as illustrated in Fig. \ref{sctld:network_dir}}. Moreover, to account for coral resistance to the disease, the intra-reef transmission function $\beta'(I_j)$ has the shape of a smooth step function of the fraction of infectious corals $I_j$ and is expressed as follows:
\begin{equation}
    \beta'(I_j) = \dfrac{\beta_0'}{2}(1+\tanh[(I_j-I_0)/\tau]),\label{eq:beta}
\end{equation}
where $I_0$ is a threshold on the infection population above which intra-reef transmission becomes significant, and $\tau$ is a measure of the interval over which the transition from low to high transmission occurs. As long as the fraction of infectious corals on sub-reef $j$ is below $I_0$, the only infection mechanism taking place is connectivity-driven transmission at rate $\beta$. Once the threshold is approached, intra-reef transmission with rate $\beta'_0$ is activated. A larger value of threshold $I_0$ corresponds to a greater resistance of corals to the disease, and therefore a slower spread of the disease within reef $j$. Coral reproduction and natural (\ie~non SCTLD-related) death rates are not taken into account in this model, which amounts to assume that they balance each other. For this study the same values were used for $\beta$ and $\beta'_0$.
%This assumption seems reasonable since coral populations of the FRT were declining prior to the emergence of the SCTLD outbreak. 
\begin{figure}
	\centering
	\includegraphics[width=.6\textwidth]{chapters/sctld/figures/network_sir.png}
	\caption{\modif{Illustration of the developed connected SIR model. The fraction of infectious corals $I_j$ on sub-reef polygon $j$ infects the fraction of susceptible individuals $S_j$ on the same sub-reef polygon if $I_j \geq I_0$. Furthermore, infectious corals on sub-reef $i$ can infect susceptible individuals of sub-reef $j$ if the two sub-reefs are connected in the connectivity matrices ($\tilde{C}_{ij} > 0$)}}.
	\label{sctld:network_dir}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=.99\linewidth]{chapters/sctld/figures/hull_example.png}
    \caption{Method used to compute the disease front displacement during a simulated time interval. \textbf{a.} Concave hull $H_0$ of the diseased sub-reefs at the beginning of the simulated period.  \textbf{b.} Concave hull $H_1$ of the sub-reefs diseased during the simulated time interval. \textbf{c.} Arrow showing the computed front displacement during simulated time interval between $H_0$ and $H_2$, the union of $H_0$ and $H_1$.}
    \label{fig:hull}
\end{figure}

### 2022-03-26 thesis.tex
modif post-defense - part I

\begin{document}
\frontmatter
  \input{include/title}
  \clearemptydoublepage
  \input{include/abstract}
  \clearemptydoublepage
  %\input{include/thanks}
  %\clearemptydoublepage
  \tableofcontents
  \clearemptydoublepage
  % \input{include/glossary}
  % \clearemptydoublepage
  \mainmatter 
%%%%
\mychapter{Introduction} \label{chap:intro}
 \input{chapters/intro/intro}    
%%
%%%%%% 
\mychapter{Processes driving the spread of SCTLD through the FRT} \label{chap:sctld}
\chaptermark{Processes driving the spread of SCTLD in the FRT}  
\input{chapters/sctld/sctld}
\mychapter{Transmission dynamics of SCTLD from the Marquesas and to the DRTO}\label{chap:drto}
\chaptermark{Transmission of SCTLD from the Marquesas to the DRTO}  
\input{chapters/drto/drto}
\mychapter{The onset of the SCTLD outbreak and the PortMiami Deep Dredge Project} \label{chap:onset}
\chaptermark{Onset of SCTLD and PortMiami Deep Dredge Project}  
\input{chapters/onset/onset}
\mychapter{Impacts of hurricanes on transport processes in the Florida Keys} \label{chap:irma}
\chaptermark{Wave-induced transport under hurricanes}  
\input{chapters/irma/irma}
%%%%%%%%%%%
%
%%%%%  
%%%
  \mychapter{Conclusions and perspectives} \label{chap:conclusion}
  \input{chapters/conclusions/conclusions}    
%%
%
  \clearemptydoublepage
  \bibliography{include/bibliography.bib}

>>>>

\begin{document}
\frontmatter
  \input{include/title}
  \clearemptydoublepage
  \input{include/abstract}
  \clearemptydoublepage
  %\input{include/thanks}
  %\clearemptydoublepage
  \tableofcontents
  \clearemptydoublepage
  \input{include/glossary}
  \clearemptydoublepage
  \mainmatter 
%%%%
\mychapter{Introduction} \label{chap:intro}
 \input{chapters/intro/intro}
%%%%%% 
\mychapter{Processes driving the spread of stony coral tissue loss disease through the Florida Reef Tract} \label{chap:sctld}
\chaptermark{Processes driving the spread of SCTLD in the FRT}  
\input{chapters/sctld/sctld}
\mychapter{Transmission dynamics of the disease from the Marquesas and to the Dry Tortugas}\label{chap:drto}
\chaptermark{Transmission of SCTLD from the Marquesas to the DRTO}  
\input{chapters/drto/drto}
\mychapter{The onset of the outbreak and the PortMiami Deep Dredge Project} \label{chap:onset}
\chaptermark{Onset of SCTLD and PortMiami Deep Dredge Project}  
\input{chapters/onset/onset}
\mychapter{Impacts of hurricanes on transport processes in the Florida Keys} \label{chap:irma}
\chaptermark{Wave-induced transport under hurricanes}
\input{chapters/irma/irma}
%%%%%%%%%%%
%
%%%%%  
%%%
  \mychapter{Conclusions and perspectives} \label{chap:conclusion}
  \input{chapters/conclusions/conclusions}    
%%
%
  \clearemptydoublepage
  \bibliography{include/bibliography.bib}
